Equivalence Testing and the Second Generation P-Value

### 2018-07-21 sgpv_vs_TOST.Rmd
clarify sentence

The SGPV has clear similarities with an already existing approach that compared the data to an equivalence range, known as equivalence testing (Lakens, 2017; Rogers, Howard, & Vessey, 1993). In the Two One-Sided Tests (TOST) procedure data is tested against the upper and lower bounds of the equivalence range. If both one-sided tests indicate that the presence of effects that fall outside of the equivalence range (or in other words, when a 90% CI falls completely within the equivalence range) an equivalence test can be used to support the abense of a meaningful effect. Surprisingly, Blume et al (2018) do not discuss equivalence testing in their article, despite the strong conceptual similarities. Here, we aim to examine the similarities and differences between equivalence testing using the TOST procedure and the SGPV.

@@

In the plot below *p*-values are calculated for the TOST equivalence testing procedure where a true population mean ranging from 140 to 150 is compared to the test value of 145 in a one-sample equivalence test where equivalence bounds are set to difference of -2 and +2 around the test value of 145. In other words, the equivalence range in the test contains all means between 143 and 147. For clarity, sample sizes consist of 1000000 observations. The population standard deviation is set to 500.

>>>>

The SGPV has clear similarities with an already existing approach that compared the data to an equivalence range, known as equivalence testing (Lakens, 2017; Rogers, Howard, & Vessey, 1993). In the Two One-Sided Tests (TOST) procedure data is tested against the upper and lower bounds of the equivalence range (e.g., a difference of -2 and +2. If both one-sided tests indicate that effects that fall outside of the equivalence range can be statistically rejected (or in other words, when a 90% CI falls completely within the equivalence range of -2 and +2) an equivalence test can be used to support the abense of a meaningful effect. Surprisingly, Blume et al (2018) do not discuss equivalence testing in their article, despite the strong conceptual similarities. Here, we aim to examine the similarities and differences between equivalence testing using the TOST procedure and the SGPV.

@@

In the plot below *p*-values are calculated for the TOST equivalence testing procedure where a true population mean ranging from 140 to 150 is compared to the test value of 145 in a one-sample equivalence test where equivalence bounds are set to difference of -2 and +2 around the test value of 145. In other words, the equivalence range in the test contains all means between 143 and 147. Blume et al (2018) rely on the z-distribution, while to TOST package uses the *t*-distribution (which is more accurate at smaller sample sizes). To make sure the SGPV give basically identical results, sample sizes consist of 1000000 observations (for which the t-distribution and z-distribution are basically identical). The population standard deviation is set to 500 to still give some variation in responses. Our conclusions should hold to the same extend for more realistic numbers (e.g., N = 100, SD = 1).
